Optimizing a PoS Tag Set for Dependency Parsing

### 2016-05-06 submission.tex
Tag Set Mapping

\paragraph{Tag Set Mapping}
The tag set modifications are realized through a mapping procedure that
concatenates the existing tag with one or more specified features for all
applicable tokens, i.e., tokens annotated with said tag and feature(s) in the
gold standard data.
%A given tag set modification involves the specification of
%a tag and the desired features we want to append to the existing,
%coarse-grained tag.

@@

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llllll@{}}
        \toprule
        \textbf{Data} & \textbf{Tag Set} & \textbf{MFT} & \textbf{Acc} &
        \textbf{LAS} & \textbf{UAS} \\
        \midrule
        \multirow{2}{*}{Dev}
        & Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        %Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
        & Optimized & 85.15\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
        \midrule
        \multirow{2}{*}{Test}
        & Original & \textbf{94.22\%} & \textbf{97.30\%} & 86.64\% & 90.07\% \\
        %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
        & Optimized & 88.08\% & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
        \bottomrule
    \end{tabular}
    \caption{Results of tagging and parsing with the optimized tag set,
        compared to the initial tag sets. Trained and tested using
        automatically assigned tags from TnT.}
    \label{finalresults}
\end{table}

@@

\section{Summary}
\label{sec:summary}
This paper has reported on a range of experiments with injecting more
fine-grained morphological distinctions into an existing PoS tag set, and then
empirically evaluating the effects both (intrinsically) in terms of tagging
accuracy and (extrinsically) in terms of parsing accuracy. Our experimental
results -- based on the annotations of the Norwegian Dependency Treebank using
the TnT PoS tagger and the Mate dependency parser -- show that the enriched tag
set leads to significantly improved parsing accuracy, even though PoS tagging
accuracy in isolation is reduced. The modified tag sets in our experiments are
defined as combinations of PoS tags and morphological features already
available in the gold annotations of the treebank. The best performing tag
configuration is in effect a PoS tag set optimized for dependency parsing of
Norwegian. However, by using another task for the extrinsic evaluation one can
of course optimize a tag set specifically for other downstream applications
instead, such as sentiment analysis, named entity recognition or any other task
making use of tagged data.

>>>>

\paragraph{Tag Set Mapping}
Based only on information already available in the gold annotations of the
treebank, the tag set modifications are realized through a mapping procedure that
concatenates existing tags with one or more specified morphological features for all
applicable tokens.
 %, i.e., tokens annotated with said tag and feature(s) in the
% gold standard data.
%A given tag set modification involves the specification of
%a tag and the desired features we want to append to the existing,
%coarse-grained tag.

@@

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llllll@{}}
        \toprule
        \textbf{Data} & \textbf{Tag Set} & \textbf{MFT} & \textbf{Acc} &
        \textbf{LAS} & \textbf{UAS} \\
        \midrule
        \multirow{2}{*}{Dev}
        & Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        %Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
        & Optimized & 85.15\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
        \midrule
        \multirow{2}{*}{Test}
        & Original & \textbf{94.22\%} & \textbf{97.30\%} & 86.64\% & 90.07\% \\
        %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
        & Optimized & 88.08\% & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
        \bottomrule
    \end{tabular}
    \caption{Results of tagging and parsing with the optimized tag set,
        compared to the original coarse tag set. Trained and tested using
        automatically assigned tags from TnT.}
    \label{finalresults}
\end{table}

@@

\section{Summary}
\label{sec:summary}
This paper has reported on a range of experiments with injecting more
fine-grained morphological distinctions into an existing PoS tag set, and then
empirically evaluating the effects both (intrinsically) in terms of tagging
accuracy and (extrinsically) in terms of parsing accuracy. Our experimental
results -- based on the annotations of the Norwegian Dependency Treebank and using
the TnT PoS tagger and the Mate dependency parser -- show that the enriched tag
set leads to significantly improved parsing accuracy, even though PoS tagging
accuracy in isolation is reduced. The modified tag sets in our experiments are
defined as combinations of PoS tags and morphological features already
available in the gold annotations of the treebank. The best performing tag
configuration is in effect a PoS tag set optimized for dependency parsing of
Norwegian. However, by using another task for the extrinsic evaluation one can
of course optimize a tag set specifically for other downstream applications
instead, such as sentiment analysis, named entity recognition or any other task
making use of tagged data.
